---
title: "Project report - Deep Learning for the Analysis of Remote Sensing Imagery from Nano Satellites"
author: "Katharina Hovestadt, Lia Kirsch, Jannis Fröhlking (439 599)"
date: "27 6 2021"
output: 
  html_document:
    css: style.css
    number_sections: yes
---

```{r standardize output, include=FALSE}
library(knitr)
opts_chunk$set(message=FALSE, warning=FALSE, echo=FALSE, fig.width = 7, fig.height = 7, fig.align = "center")
```

```{r env, results=F,warning=F,message=F}
library(keras)
library(tensorflow)
library(tfdatasets)
library(purrr)
library(ggplot2)
library(rsample)
library(stars)
library(terra)
library(raster)
library(reticulate)
library(mapview)
```

# Introduction

# Data

What you need:
- planet 3 band images of train region (Bahrain)
- planet 3 band images of study region (Qatar)
- osm shapefile containing streets of Bahrain @see: [Link to OSM download]

## Train data

### Preprocess satellite images
```{r function to preprocess satellite tiles}
setNAs <- function(test_tile){
  # set zero values to NA
  test_tile[test_tile[]==0] <- NA
  # crop to non NA values
  test_tile = trim(test_tile)
  # remove fourth band (NIR)
  test_tile = dropLayer(test_tile,4)
  return (test_tile)
  }
```

```{r merge single raster images}
# Set path to folder containing Planet satellite images of Bahrain
if(!file.exists("training_unet/bahrain_2021.tif")){
  setwd("Bahrain_2021_3Band/files")
  files = list.files(pattern = ".tif$") 
  filesList = as.list(files)
  stackList = lapply(filesList, stack)
  stackList = lapply(stackList, setNAs)
  mosaic_bahrain = mosaic(stackList[[1]],stackList[[2]],stackList[[3]], stackList[[4]], fun = mean)
  setwd("~/")
  writeRaster(mosaic_bahrain, "training_unet/bahrain_2021.tif")
}


if(!file.exists("testarea_unet/doha_2021.tif")){
  setwd("2021/files")
  files = list.files(pattern = ".tif$") 
  filesList = as.list(files)
  stackList = lapply(filesList, stack)
  stackList = lapply(stackList, setNAs)
  mosaic_bahrain = mosaic(stackList[[1]],stackList[[2]],fun = mean)
  setwd("~/")
  writeRaster(mosaic_bahrain, "testarea_unet/doha_2021.tif")
}

if(!file.exists("testarea_unet/doha_2019.tif")){
  setwd("2019/files")
  files = list.files(pattern = ".tif$") 
  filesList = as.list(files)
  stackList = lapply(filesList, stack)
  stackList = lapply(stackList, setNAs)
  mosaic_bahrain = mosaic(stackList[[1]],stackList[[2]],fun = mean)
  setwd("~/")
  writeRaster(mosaic_bahrain, "testarea_unet/doha_2019.tif")
}

if(!file.exists("testarea_unet/doha_2017.tif")){
  setwd("2017/files")
  files = list.files(pattern = ".tif$") 
  filesList = as.list(files)
  stackList = lapply(filesList, stack)
  stackList = lapply(stackList, setNAs)
  mosaic_bahrain = mosaic(stackList[[1]],stackList[[2]],fun = mean)
  setwd("~/")
  writeRaster(mosaic_bahrain, "testarea_unet/doha_2017.tif")
}
```

### Preprocess street shapefile
```{r crop street shapefile to satellite image}
# load planet image of train area
st_bahrain = stack("training_unet/bahrain_2021.tif")
# load openstreetmap data of the streegts
streets_all = read_sf("data/shp/streets_osm.shp")
# transform osm data to crs of bahrain image
streets_utm = st_transform(streets_all, crs = crs(st_bahrain))
streets = st_crop(streets_utm, st_bbox(st_bahrain))
```


#### Rasterize streets shapefile
```{r rasterization of all roads}
# creates a SpatVector object
v_lines = vect(streets)
# prere the output raster size
r = rast(v_lines, ncol = ncol(st_bahrain), nrow = nrow(st_bahrain))
x <- rasterize(v_lines, r, touches = TRUE, background = 0)
par(mfrow=c(1,2))
plot(x, col = c("white","black"), axes = FALSE, legend = FALSE)
plotRGB(st_bahrain, r = 1, g=2, b=3)
```


Buffer streets to get better rasterization result, because main roads have a higher width
```{r buffer}
primaries = streets[streets$fclass == "primary",]
secondaries = streets[streets$fclass == "secondary",]
tertiaries = streets[streets$fclass == "tertiary",]
trunks = streets[streets$fclass == "trunk",]

poly_streets = rbind(primaries,secondaries, tertiaries, trunks)

st_geometry(poly_streets[poly_streets$fclass == "trunk",]) = st_geometry(st_buffer(trunks, dist = 6))
st_geometry(poly_streets[poly_streets$fclass == "primary",]) = st_geometry(st_buffer(primaries, dist = 5))
st_geometry(poly_streets[poly_streets$fclass == "secondary",]) = st_geometry(st_buffer(secondaries, dist = 4))
st_geometry(poly_streets[poly_streets$fclass == "tertiary",]) = st_geometry(st_buffer(tertiaries, dist = 3))
```

```{r rasterization of main roads}
v_poly = vect(poly_streets)
r_poly = rast(v_poly, ncol = ncol(st_bahrain), nrow = nrow(st_bahrain), extent = ext(x))
x_poly <- rasterize(v_poly, r_poly, touches = TRUE, background = 0)
par(mfrow=c(1,2))
plot(x_poly, col =c("white","black"), axes = FALSE, legend = FALSE)
plot(x, col =c("white","black"), axes = FALSE, legend = FALSE)
plotRGB(st_bahrain, r = 1, g=2, b=3)
```

Combine resulting two raster layers
```{r combine resulting rasters}
result = max(x_poly,x)
writeRaster(result, "training_unet/street_mask.tif", overwrite = TRUE)
```

### Create subsets for the unet training
```{r, dl_subsets}
dl_subsets <- function(inputrst, targetsize, targetdir, targetname="", img_info_only = FALSE, is_mask = FALSE){
  require(jpeg)
  require(raster)
  
  #determine next number of quadrats in x and y direction, by simple rounding
  targetsizeX <- targetsize[1]
  targetsizeY <- targetsize[2]
  inputX <- ncol(inputrst)
  inputY <- nrow(inputrst)
  
  #determine dimensions of raster so that 
  #it can be split by whole number of subsets (by shrinking it)
  while(inputX%%targetsizeX!=0){
    inputX = inputX-1  
  }
  while(inputY%%targetsizeY!=0){
    inputY = inputY-1    
  }
  
  #determine difference
  diffX <- ncol(inputrst)-inputX
  diffY <- nrow(inputrst)-inputY
  
  #determine new dimensions of raster and crop, 
  #cutting evenly on all sides if possible
  newXmin <- floor(diffX/2)
  newXmax <- ncol(inputrst)-ceiling(diffX/2)-1
  newYmin <- floor(diffY/2)
  newYmax <- nrow(inputrst)-ceiling(diffY/2)-1
  rst_cropped <- suppressMessages(crop(inputrst, extent(inputrst,newYmin,newYmax,newXmin,newXmax)))
  #writeRaster(rst_cropped,filename = target_dir_crop)
  
    #return (list(ssizeX = ssizeX, ssizeY = ssizeY, nsx = nsx, nsy =nsy))
    agg <- suppressMessages(aggregate(rst_cropped[[1]],c(targetsizeX,targetsizeY)))
    agg[]    <- suppressMessages(1:ncell(agg))
    agg_poly <- suppressMessages(rasterToPolygons(agg))
    names(agg_poly) <- "polis"
    
    pb <- txtProgressBar(min = 0, max = ncell(agg), style = 3)
    for(i in 1:ncell(agg)) {
      
      # rasterOptions(tmpdir=tmpdir)
      setTxtProgressBar(pb, i)
      e1  <- extent(agg_poly[agg_poly$polis==i,])
      
      subs <- suppressMessages(crop(rst_cropped,e1))
      #rescale to 0-1, for jpeg export
      if(is_mask==FALSE){
        
        subs <- suppressMessages((subs-cellStats(subs,"min"))/(cellStats(subs,"max")-cellStats(subs,"min")))
      } 
      #write jpg
      
     
      writeJPEG(as.array(subs),target = paste0(targetdir,targetname,i,".jpg"),quality = 1)
      
      #writeRaster(subs,filename=paste0(targetdir,"SplitRas_",i,".tif"),overwrite=TRUE) 
      #return(c(extent(rst_cropped),crs(rst_cropped)))
    }
    close(pb)
    #img_info <- list("tiles_rows"=nrow(rst_cropped)/targetsizeY, "tiles_cols"=ncol(rst_cropped)/targetsizeX,"crs"= crs(rst_cropped),"extent"=extent(rst_cropped))
    #writeRaster(rst_cropped,filename = paste0(targetdir,"input_rst_cropped.tif"))
    rm(subs,agg,agg_poly)
    gc()
    return(rst_cropped)
  
}
```

```{r create training subsets}
inputrst = stack("training_unet/bahrain_2021.tif")
inputrst_mask = raster("training_unet/street_mask.tif")
plotRGB(inputrst, r = 1, g=2, b=3)
plot(inputrst_mask, col=c("white","black"))
# Create subsets of Bahrain and Street mask
#dl_subsets(inputrst = inputrst, targetsize = c(448,448), targetdir = "training_unet/imgs/")
#dl_subsets(inputrst = inputrst_mask, targetsize = c(448,448), targetdir = "training_unet/masks/")
```

## Test area
```{r mosaic testarea}
if(file.exists("testarea_unet/doha_2021.tif")){
  plotRGB(stack("testarea_unet/doha_2021.tif"),r=1,b=3)
}else{
  test_files = list.files("2021/files", pattern = ".tif$") 
  # put tiles into a list of raster stacks
  setwd("2021/files")
  stackList_test = lapply(test_files, stack)
  
  # test one image
  # testImage = stackList_test[[1]]
  # testImage = crop(testImage, e)
  
  stackList_test = lapply(stackList_test, setNAs)
  mosaic_doha = mosaic(stackList_test[[1]],
                       stackList_test[[2]],
                       stackList_test[[3]],
                       fun = mean)
  # r = stackList_test[[1]]
  # r = crop(r, extent(532500,540000,2783000, 2784000)) # currently used for testing
  setwd("~/")
  writeRaster(mosaic_doha, "testarea_unet/doha_2021.tif",overwrite = TRUE)
}
```


```{r create testarea subsets}
#qatar_testarea = stack("2021/files/20170422_062109_1030_3B_Visual_clip.tif")
qatar_testarea = stack("testarea_unet/doha_2019.tif")

# Crashes the r Session
raster_cropped <- dl_subsets(inputrst = qatar_testarea, targetsize = c(448,448), targetdir = "testarea_unet/2019_subsets/")

print("Testarea subset creation done!")
```

# Methods


# Results


# Discussion/Conclusion


# References